---
title: "Consulting Memo"
author: "Carolyn Clayton"
date: "Wednesday, February 25, 2015"
output: pdf_document
graphics: yes
---
``` {r packages, include=FALSE}
library(ggplot2)
library(psych)
library(plyr)
```

```{r data, eval=TRUE, echo=FALSE, results='hide'}
setwd(dir = "C:/Users/Carolyn/Dropbox/ASchool/! Consulting II/Homework 6")

## Inputting the data

Alfuzosin <- c(1, 1, 2, 2, 3, 1, 0, 3, 3, 2, 5, 2, 1, 3, 2, 1, 1, 2, 1, 2, 2, 2, 2, 1, 1, 5, 1, 2, 3, 1, 2, 5, 1, 5, 3, 3, 3, 2, 5, 2, 3, 4, 3, 4, 5, 3, 4, 3, 2, 3, 5, 2, 2, 3)
AfterAlfuzosin <- c(10, 7, 4, 4, 3, 5, 6, 7, 7, 3, 5, 2, 4, 4, 6, 3, 6, 2, 5, 3, 5, 2, 4, 2, 3, 6, 4, 5, 4, 7, 4, 3, 3, 4, 2, 4, 4, 3)
Drug <- vector()
i <- 1
repeat{ if(i <= length(Alfuzosin))
  {Drug[i] = 0}
    else if(i <= length(Alfuzosin) + length(AfterAlfuzosin))
  {Drug[i] = 1}
  i <- i+1
  if(i > length(Alfuzosin) + length(AfterAlfuzosin)) break()
}
Drug <- factor(Drug, 
               levels = c(0, 1), 
               labels = c("Alfuzosin", "AfterAlfuzosin"))
Urination <- c(Alfuzosin, AfterAlfuzosin)
Days <- 1:(length(Alfuzosin) + length(AfterAlfuzosin))
UrinationData <- data.frame(Days, Urination, Drug)

# Create fortnightly blocks of the data
fortnight <- c(1,2,3,4,5,6,7)
fortnight <- rep(fortnight, each=14, length.out = 92)
UrinationData <- cbind(UrinationData, fortnight)  # Add fortnight to the dataframe.

# Compute means by each fortnight.
Means <- by(UrinationData$Urination, UrinationData$fortnight, mean)
Means <- rep(Means, each=14, length.out=92)  # Create a vector of the means so there's one data point for each row in the dataframe.
UrinationData <- cbind(UrinationData, Means)  # Add Means to the dataframe

# Descriptives
descriptives <- describeBy(UrinationData$Urination, group = Drug)

```

```{r tests, eval=TRUE, echo=FALSE, results='hide'}
# Tests

ftest <- var.test(UrinationData$Urination ~ UrinationData$Drug)
ttest <- t.test(UrinationData$Urination ~ UrinationData$Drug, alternative = "two.sided", paired = F, conf.level = 0.95, var.equal = FALSE)

# Performing nonparametric tests just in case data are not normal. 
wilcoxtest <- wilcox.test(UrinationData$Urination ~ UrinationData$Drug)

# Code to automatically input a < 0.001 if the p-value is less than 0.001, rather than inputting an = 0.001 or = 0.000 (depending on rounding)
x <- 0
p.value <- function(x) {
    if(x < 0.001) {
      x <- "< 0.001"
    } else {
      x <- paste("= ", round(x, digits=3), sep = "")
    }
  return(x)
}
```

Dear Sir:

Below is a summary of nightly urinations on Alfuzosin and off Alfuzosin:

```{r summary,results='asis',echo=FALSE}
# Descriptives table
knitr::kable(ddply(UrinationData, ~Drug, summarise, mean=round(mean(Urination), digits = 3),sd=round(sd(Urination), digits = 3)), caption = "Nightly Urinations On and Off Alfuzosin")
```

In the analysis of the data, an F test was run to compare the variances of the two groups. This was determined to be significant, F($`r ftest$parameter[1]`$, $`r ftest$parameter[2]`$) = $`r round(ftest$statistic, digits = 3)`$, p $`r p.value(ftest$p.value)`$, 95% CI($`r round(ftest$conf.int[1], digits = 3)`$, $`r round(ftest$conf.int[2], digits = 3)`$); therefore a t-test for unequal variances was conducted. The result of this t-test was that there is a significant difference in average nightly urinations between nights on Alfuzosin and off Alfuzosin, t($`r round(ttest$parameter, digits = 3)`$) = $`r round(ttest$statistic, digits = 3)`$, p $`r p.value(ttest$p.value)`$, 95% CI($`r round(ttest$conf.int[1], digits = 3)`$, $`r round(ttest$conf.int[2], digits = 3)`$). The mean nightly urinations on Alfuzosin was $`r round(descriptives$Alfuzosin$mean, digits = 3)`$ (SD = $`r round(descriptives$Alfuzosin$sd, digits = 3)`$), while the nightly urinations off Alfuzosin was $`r round(descriptives$AfterAlfuzosin$mean, digits = 3)`$ (SD = $`r round(descriptives$AfterAlfuzosin$sd, digits = 3)`$). 

We have also plotted the nightly urinations over time, and have taken fortnightly means (as depicted in the graph below):

```{r plot1, echo=FALSE, results='hide', fig.width = 4, fig.height = 3, fig.align = 'center', dev='pdf'}
# Note: There was an issue with centering this in the PDF. Issue fixed by info contained on https://github.com/rstudio/rmarkdown/issues/325 . Added graphics: yes to YAML header.
g <- ggplot(UrinationData, aes(x = Days, y = Urination))  # Specify the data
g + geom_point(shape = 1, size = 2) + # Add data points
  geom_segment(aes(x = seq_along(Days), y = Means, xend=(seq_along(Days) + 1), yend = Means)) +  # Create the mean lines.
  geom_vline(xintercept = 54.5, size = 1) +  # Add vertical line
  theme_bw(base_size = 10) +  # Change theme, font, and font size
  labs(x = "Day (May-Aug 2006)") +  # Change x label
  labs(y = "# void") +  # Change y label
  annotate("text", x = 25, y = 9, label = "On Alfuzosin", size = 4) +  # Add On Alfuzosin annotation to plot
  annotate("text", x = 75, y = 9, label = "Off Alfuzosin", size = 4) # Add On Alfuzosin annotation to plot 
```
\begin{center}Plot 1: Nightly Urinations with Fortnightly Means\end{center}
  
Given these results, we would conclude that alfuzosin may be efficacious for you; however, there was a reduction in nightly urinations immediately prior to switching to Alfuzosin. Whether the reduction in night time urinations on Alfuzosin is sufficient to warrant the potential side effects is left up to your discretion.    

Warmly,  
Carolyn

