---
title: "Consulting Memo"
author: "Carolyn Clayton"
date: "Tuesday, January 20, 2015"
output: html_document
---
``` {r load_packages, include = FALSE}
library(psych)
library(ggplot2)
```
This is a sample consulting memo originally created to partially fulfill requirements for Consulting II at the University of Colorado Denver.  

```{r echo = FALSE, results='hide'}
## Inputting the data

Alfuzosin <- c(1, 1, 2, 2, 3, 1, 0, 3, 3, 2, 5, 2, 1, 3, 2, 1, 1, 2, 1, 2, 2, 2, 2, 1, 1, 5, 1, 2, 3, 1, 2, 5, 1, 5, 3, 3, 3, 2, 5, 2, 3, 4, 3, 4, 5, 3, 4, 3, 2, 3, 5, 2, 2, 3)
AfterAlfuzosin <- c(10, 7, 4, 4, 3, 5, 6, 7, 7, 3, 5, 2, 4, 4, 6, 3, 6, 2, 5, 3, 5, 2, 4, 2, 3, 6, 4, 5, 4, 7, 4, 3, 3, 4, 2, 4, 4, 3)
Drug <- vector()
i <- 1
repeat{ if(i <= length(Alfuzosin))
  {Drug[i] = 0}
    else if(i <= length(Alfuzosin) + length(AfterAlfuzosin))
  {Drug[i] = 1}
  i <- i+1
  if(i > length(Alfuzosin) + length(AfterAlfuzosin)) break()
}
Drug <- factor(Drug, 
               levels = c(0, 1), 
               labels = c("Alfuzosin", "AfterAlfuzosin"))
Urination <- c(Alfuzosin, AfterAlfuzosin)
UrinationData <- data.frame(Urination, Drug)
```

Dear Sir:

I have performed a two-sample t-test analysis on your data. Below you will find a boxplot of nightly urinations by drug status. 

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = FALSE, results = 'hide'}
ggplot(UrinationData, aes(x = Drug, y = Urination, fill = Drug)) +
  geom_boxplot(alpha = 0.5) + 
  stat_summary(fun.y = mean, geom = "point", shape = 5) +
  guides(fill = FALSE) +
  ggtitle("Boxplot of Nightly Urinations by Drug Status")
dev.off() 
```
  
```{r echo = FALSE, results = 'hide'}
options(digits = 3)
# Descriptives
descriptives <- describeBy(UrinationData$Urination, group = Drug)

# Tests

f.test <- var.test(UrinationData$Urination ~ UrinationData$Drug)
t.test <- t.test(UrinationData$Urination ~ UrinationData$Drug, alternative = "two.sided", paired = F, conf.level = 0.95, var.equal = FALSE)

x <- 0
p.value <- function(x) {
    if(x < 0.001) {
      x <- "< 0.001"
    } else {
      x <- cat("= ", x, sep = "")
  }
  return(x)
}

```
In the analysis of the data, an F test was run to compare the variances of the two groups. This was determined to be significant, F(`r f.test$parameter[1]`, `r f.test$parameter[2]`) = `r f.test$statistic`, p = `r f.test$p.value`, 95% CI[`r f.test$conf.int[1]`, `r f.test$conf.int[2]`]; therefore a t-test for unequal variances was conducted. The result of this t-test was that there is a significant difference in average nightly urinations between nights on Alfuzosin and off Alfuzosin (t(`r t.test$parameter`) = `r t.test$statistic`, p `r p.value(t.test$p.value)`, 95% CI[`r t.test$conf.int[1]`, `r t.test$conf.int[2]`]). The mean nightly urinations on Alfuzosin was `r descriptives$Alfuzosin$mean` (SD = `r descriptives$Alfuzosin$sd`), while the nightly urinations off Alfuzosin was `r descriptives$AfterAlfuzosin$mean` (SD = `r descriptives$AfterAlfuzosin$sd`). 

Given these results, we would conclude that alfuzosin is efficacious; however, whether the reduction in night time urinations is sufficient to warrant the potential side effects is left up to your discretion.  

Warmly,  
Carolyn

